Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-11 paper.Rmd
revert lists to markdown style

B\&S claim (page 3) to show that a model informed by procedural theories of escalation and de-escalation can predict the onset of civil wars remarkably accurately. Indeed they argue that this so-called theoretical model outperforms four other "more mechanical" alternatives. Second, they claim that the integration of structure with process is better over short forecasting windows. Third, they preregistered the list of thirty countries that have the highest risk of civil war onset. They claim that such prospective predictions are rare in the literature when, in fact, they have been routine for many years with several prominent projects. B\&S claim to be unique in assessing these forecasts. A qualitative analysis of their predictions allows them to conclude that their model is robust, and a more precise test is undertaken. We will return to their analysis later, after correcting the procedural mistakes we found in their research process.

@@

\begin{itemize}
\item Escalation: a set of ten indicators, putatively drawn from a theoretical escalation model, for interactions between the government on one side and opposition or rebel actors on the other. 
\item Quad: ICEWS quad counts, i.e. material conflict, material cooperation, verbal conflict, verbal cooperation, for interactions between the government and opposition or rebels. These are directed, thus making for four directed dyads, which with four quad categories make for a total of sixteen covariates. 
\item Goldstein: -10 (conflictual) to 10 (cooperative) scores derived from the ICEWS data for the same four directed dyads, for a toal of four covariates. 
\item CAMEO: counts for all CAMEO event codes over the four actor dyads, totaling $1,160$ covariates, which are mostly zero for any country in any month. 
\item Average: unweighted average of the predictions from the four models briefly described above.
\end{itemize}

@@

\begin{itemize}
\item The theory-driven escalation model is outperformed both by the low-effort 1,160 predictor all-CAMEO model.
\item The Average ensemble model and the CAMEO models outperform the escalation model in all instances.
\item Adding structural variables substantially improves the escalation model's performance.
\end{itemize}

>>>>

B\&S claim (page 3) to show that a model informed by procedural theories of escalation and de-escalation can predict the onset of civil wars "remarkably accurately". Indeed they argue that this so-called theoretical model outperforms four other "more mechanical" alternatives. Second, they claim that the integration of structure with process is better over short forecasting windows. Third, they preregistered the list of thirty countries that have the highest risk of civil war onset. They claim that such prospective predictions are rare in the literature when, in fact, they have been routine for many years with several prominent projects. B\&S claim to be unique in assessing these forecasts. A qualitative analysis of their predictions allows them to conclude that their model is robust, and a more precise test is undertaken. We will return to their analysis later, after correcting the procedural mistakes we found in their research process.

@@

- Escalation: a set of ten indicators, putatively drawn from a theoretical escalation model, for interactions between the government on one side and opposition or rebel actors on the other. 
- Quad: ICEWS quad counts, i.e. material conflict, material cooperation, verbal conflict, verbal cooperation, for interactions between the government and opposition or rebels. These are directed, thus making for four directed dyads, which with four quad categories make for a total of sixteen covariates. 
- Goldstein: -10 (conflictual) to 10 (cooperative) scores derived from the ICEWS data for the same four directed dyads, for a toal of four covariates. 
- CAMEO: counts for all CAMEO event codes over the four actor dyads, totaling $1,160$ covariates, which are mostly zero for any country in any month. 
- Average: unweighted average of the predictions from the four models briefly described above.

@@

- The theory-driven escalation model is outperformed both by the low-effort 1,160 predictor all-CAMEO model.
- The Average ensemble model and the CAMEO models outperform the escalation model in all instances.
- Adding structural variables substantially improves the escalation model's performance.
